Meta-analytic approaches and new effect sizes for treatment effects in comparative physiology

### 2021-05-17 MS/ms.Rmd
Adding some refs, just getting ideas on paper

---
title: "Meta-analytic approaches and new effect sizes to account for treatment differences across studies in comparative physiology"
author: Daniel W.A. Noble, Patrice Potter, Sammy Burke, Szymek Drobniak, Malgorzata Lagisz, Rose O’Dea, Shinichi Nakagawa
date: "`r Sys.Date()`"
bibliography: ../bib/refs.bib
csl: ../bib/the-journal-of-experimental-biology.csl
output: 
  bookdown::word_document2:
    toc: no
    toc_depth: 6
    number_sections: false
    reference_docx: template.docx
  bookdown::html_document2:
    code_folding: hide
    number_sections: no
    toc: yes
    toc_depth: 6
    toc_float: yes
    bibliography: refs.bib
editor_options: 
  chunk_output_type: console
---

@@

# **Abstract**

@@

>>>>

---
title: "Meta-analytic approaches and new effect sizes for treatment effects in comparative physiology"
author: Daniel W.A. Noble, Patrice Potter, Sammy Burke, Szymek Drobniak, Malgorzata Lagisz, Rose O’Dea, Shinichi Nakagawa
date: "`r Sys.Date()`"
bibliography: ../bib/refs.bib
csl: ../bib/the-journal-of-experimental-biology.csl
output: 
  bookdown::word_document2:
    toc: no
    toc_depth: 6
    number_sections: false
    reference_docx: template.docx
  bookdown::html_document2:
    code_folding: hide
    number_sections: no
    toc: yes
    toc_depth: 6
    toc_float: yes
    bibliography: refs.bib
editor_options: 
  chunk_output_type: console
---

@@

## **Abstract**

@@

## **Introduction**
Meta-analysis provides unbiased and quantitatively informed answers to research questions in comparative physiology by using a transparent and reproducible search strategy to synthesise research results across studies and identify sources of variation [@Nakagawa2017; @Gurevitch2018; @Cooper2009; @Arnqvist1995; @Koricheva2013]. Meta-analysing independent studies provides greater statistical power and precision than what any individual study on its own would be able to provide on a given question - particularly given that most studies are already under-powered in ecological and evolutionary research [@Forstmeier2017; @Jennions2003]. By expressing study effects on a common scale (i.e., effect size) a meta-analyst can gain broader insight into the direction and efficacy of a particular treatment effect or strength of a relationship between two variables of interest [@Gurevitch2018; @Koricheva2013].  A critically important feature of meta-analaysis that provides improved precision when estimating overall effects is the ability to account for sampling error - this can be derived assuming the sampling distribution of a given effect size is known [].

Despite the widespread adoption and well established methodological procedures of meta-analysis it has been criticised for mixing 'apples and oranges'; in otherwords, mixing effects from studies that lack comparability [@Arnqvist1995; @Gallo1979; @Carpenter2020; @Stewart2010]. Lack of comparability could result from studies differing in experimental design, species, measurement variables and sampling units, which can impact the generalisability of meta-analytic results [@Arnqvist1995; @Stewart2010]. Indeed, study specific differences in comparative physiology are widspread. For example, experimental designs can vary in the temperatures that they apply or in the dosages of chemical modifiers used in experimental treatments [e.g., ]. They often inlcude a diversity of study species.

The goal of our review is to reassure comparative physiologists that heterogeneity among effect sizes can be comfortably dealt with through modifications of existing effect sizes measures and using multilevel meta-regression models. We first conduct a survey of published meta-analyses in the field of comparative physiology to gauge methodological and meta-analytic practices - assessing what types of effect sizes are commonly used, how often effect size estimates are impacted by continuous treatment differences and whether meta-analytic models (i.e., models that account for sampling variance) are commonly applied. We then introduce new effect sizes and their assocaited sampling variance to provide comparative physiologists opportunities to explictly incorperate treatment differences into the effect size estimates. We then describe how more complex treatment differences, such as dosage differences which may have more complex non-linear effects, can be accomodated using multilevel meta-regression models. Expanding the meta-analtic toolkit for comparative physiologists will provide new opportunities to address global scale problems that attempt to undersatnd how organisms will cope with rapidly changing environments in the future.

## **Meta-Analytic Models that Control for Treatment Differences Within and Across Studies**

Quantifying effect size heterogenity (i.e., effect size variance) provides a pathway to explore hypotheses thought to explain why effect might vary across studies.

## **References**
